\documentclass[conference]{IEEEtran}
%\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
%% \usepackage{cite}
\usepackage{hyperref}
\usepackage{svg}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{amsthm} % Allows usage of '\theoremstyle'

\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage[font=small,labelfont=bf]{caption} % small fontsize in caption

\usepackage{siunitx}

%% \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
%%     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\usepackage[
    backend=biber,
    style=ieee,
    sorting=none,
    %% citestyle=authoryear
    ]{biblatex}
\addbibresource{bibliography.bib}

\usepackage[normalem]{ulem}


\begin{document}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\vecs}[1]{\boldsymbol{#1}}
\newcommand{\matr}[1]{\boldsymbol{#1}}
\newcommand{\matd}[1]{\mathcal{#1}}

\newcommand{\dotprod}[2]{\left\langle {#1}, \, {#2} \right\rangle}
\newcommand{\normdotprod}[2]{\frac{\left\langle #1, \, #2 \right\rangle}{\| #1 \| \, \| #2 \|}}


\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{lemma}{Lemma}[section]
% \theoremstyle{definition}
\newtheorem{definition}{Definition}[section]


\title{
Obstacle Aware Passive Control \\
to Follow Desired Velocities
}
\author{
  \IEEEauthorblockN{Lukas Huber, Trinca Thibaud, Jean-Jacques Slotine, Aude Billard}
}

\maketitle
\thispagestyle{plain}
\pagestyle{plain}

\begin{abstract}
Position-based velocity evaluation has revolutionized robot control methodologies, endowing the systems with enhanced resilience against noise and disturbances. The desired velocities can incorporate obstacle avoidance capabilities, further bolstered by passive torque control to ensure safe responses to external perturbations. 
In this paper, we propose a novel method for designing passivity controllers that internalizes the presence of obstacles during the handling of external forces.

Our method introduces a compliant behavior away from obstacles while strategically increasing control damping upon approaching obstacles, thereby enabling the robot to maintain heightened environmental awareness. The algorithm's passivity is proven to remain stable across all scenarios, effectively countering disturbances. Additionally, we show superior behavior in collision avoidance and overall stability, particularly in scenarios with high damping values. 

We additionally observe that the obstacle-aware passive controller exhibits superior tracking and collision avoidance performance in the presence of noise.
The practicality and effectiveness of our approach are validated through implementation on a 7DoF robot arm. Comparative analysis against a baseline method showcases the algorithm's prowess in achieving superior collision rejection, underscoring its potential for real-world applications. Overall, our proposed obstacle-aware passivity controller represents a significant step towards enhanced robot control and safety in complex and uncertain environments.
% Position-based dynamical systems have enabled new ways of robot control, making the system more robust to noise or disturbances.
% The dynamics can be designed to incorporate obstacle avoidance capabilities, and combined with passive torque control to safely react to disturbances.
% In this work, we propose a novel method to design the passivity controller to improve the handling of external forces. 
% The method maintains compliant behavior away from obstacles but increases the damping of the control when approaching one of them, allowing the robot to be aware of its environment.
% The algorithm is shown to remain passive (stable) with respect to disturbances in all scenarios. 
% Moreover, we obtained improved collision avoidance, and stability in the presence of high damping values. 
% The obtacle-aware passive controller shows improved tracking and collision avoidance performance in the presence of noise. 
% Finally, the algorithm was implemented on a 7DoF robot arm. Superior collision rejection was achieved compared to the baseline method.
\end{abstract}

\begin{IEEEkeywords}
Obstacle avoidance, dynamical systems, passivity, human-robot interaction
\end{IEEEkeywords}

\input{input/introduction}
\input{input/preliminaries}

\subsection{Problem Statement}
Following assumptions are made about the desired velocity $\vecs f(\vecs\xi)$:
\begin{enumerate}
    \item $\vecs f(\vecs\xi)$ is continuously  for all reachable states.
    %% \item $\vecs f(\vecs\xi)$ has a single equilibrium point $\vecs\xi^a$, such that $\{ \vecs \xi : \vecs f(\vecs \xi) = \vecs 0 \} = \{ \vecs\xi^a \}$. 
    \item $\vecs f(\vecs\xi)$ is bounded, i.e., there exists a constant $v^{\mathrm{max}} \in \mathbb{R}$ such that $\| \vecs f(\vecs\xi) \| \leq v^{\mathrm{max}} \;\; \forall \, \vecs \xi \in \mathbb{R}^N$
    \item It leads to a collision-free motion $\vecs{n_o}(\vecs\xi)^T \vecs f(\vecs\xi) \geq 0$ as $\Gamma_o \rightarrow 1 
  \quad \forall o = 1 .. N^{\mathrm{obs}}$
  with the normal $\vecs n_o (\vecs \xi)$ and distance $d_o$ of the $o$-th obstacle. 
\end{enumerate}

Note that velocities obtained using DSM, given in \eqref{eq:modulated}, fulfill these conditions if velocity function $\vecs f^I$ is continuous and bounded.

\input{input/obstacle_aware_passivity}
\input{input/passivity_proof}
\input{input/collision_avoidance}
\input{input/discrete_time}

\input{input/results}

% \appendix
% \section{Appendix}
% \label{Appendix}

\renewcommand*{\bibfont}{\footnotesize}
\printbibliography

\end{document}
